---
title: "Statistics using R"
output: html_notebook
---

```{r}
install.packages("prob")
```
```{r}
library(prob)
library(tidyverse)
```
```{r}
s_two_dice <- rolldie(2)
View(s_two_dice)
nrow(s_two_dice)
```
```{r}
s_two_dice <- s_two_dice %>%
  mutate(total = X1 + X2)

s_two_dice %>%
  filter(total > 10)
```
```{r}
first_die_even <- s_two_dice %>%
  filter(X1 %% 2 == 0) %>%
  filter(total < 10)
first_die_even
```
# Counting (we need to know how many total outcomes there are)
P(enent) = num(outcomes)/num(all_outcomes)

```{r}
s_two_dice <- rolldie(2)

size_s <- nrow(s_two_dice) # [36] outcomes

total_seven <- s_two_dice %>%
  filter(X1 + X2 == 7)

size_seven <- nrow(total_seven) # [6] outcomes in total that return 7.
prob_total_seven <- size_seven/size_s # probability of this occuring is 0.1666667.
```
## combining odds

```{r}
s_die <- rolldie(1)
s_coin <- tosscoin(1)

s_total <- merge(s_die, s_coin)
```
# Permutations (specific numbers in specific order)

number of permutations (nPr) = N! / (N-r)!
4! = 24
```{r}
install.packages("gtools")
library(gtools)
```
```{r}
employees <- c("Alice", "Bob", "Carole", "David", "Elaine", "Fred", "Ghislaine", "Hua", "Inigo", "Hugo")
```
```{r}
senior_staff <- as_tibble(permutations(n = 10,
                             r = 3,
                             v = employees,
                             repeats.allowed = FALSE)
)
```
```{r}
senior_staff <- senior_staff %>%
  rename(CEO = V1, CTO = V2, COO = V3)
```
What is the prob of Alice being CEO and Bob being CTO?
```{r}
alice_and_bob <- senior_staff %>%
  filter(CEO == "Alice", CTO == "Bob")
prob_alice_and_bob <- nrow(alice_and_bob)/nrow(senior_staff)
```
```{r}
letter_combinations <- c("A", "B", "C", "D", "E")
perm_letters_repeats <- as_tibble(permutations(n = 5,
                                       r = 3,
                                       v = letter_combinations,
                                       repeats.allowed = TRUE))

perm_letters_no_repeats <- as_tibble(permutations(n = 5,
                                       r = 3,
                                       v = letter_combinations,
                                       repeats.allowed = FALSE))
```
# Combinations. Certain numbers, but the order does not matter.
NCr = N!/(r!*(N-r)!)

taking 5 card hand out of 52 pack: N = 52, r = 5
```{r}
card_draws <- choose(52,5)
```
```{r}
colours <- c("red", "green", "yellow")
three_colour_permutations <- permutations(n= 3,
                                          r = 3,
                                          v = colours,
                                          repeats.allowed = FALSE)

three_colour_combinations <- combinations(n = 3,
                                          r = 3,
                                          v = colours)

```
```{r}
lottery_combinations <- choose(50, 6) # total combinations
```
```{r}
lottery_probability <- 10 / lottery_combinations
```
# Probability combinations. Complement

P(Ac) = 1 - P(A)

probability of rolling 6 on a dice is 1/6, probability of rolling anything but 6 is 5/6.

```{r}
s_four_coins <- tosscoin(4) %>%
  mutate(number_of_heads = as.numeric(toss1 == "H") + 
           as.numeric(toss2 == "H") + 
           as.numeric(toss3 == "H") + 
           as.numeric(toss4 == "H"),
         number_of_tails = as.numeric(toss1 == "T") +
           as.numeric(toss2 == "T") +
           as.numeric(toss3 == "T") +
           as.numeric(toss4 == "T") 
          ) %>%
  filter(number_of_heads >= 1 & number_of_tails >= 1)

```

# Intersection
Probability of two things happening at the same time.

P(A & B) 
```{r}
rolldie(2) %>%
  filter(X1%%2 == 0 & X2 == 5)
```


# Union. 
Probabililty of A or B happening. 


P(A or B)
```{r}
# probability of first die is even and second is 5.
rolldie(2) %>%
  filter(X1%%2 == 0 | X2 == 5) # 21 chances, ot total 36
20/52
```

# Conditional probability

P(A|B) probability of A given that B has happened = P(A intersection B)/P(B)

# Independent Events.


