---
title: "Date/Time"
output: html_notebook
---

Library to use: "lubridate" !!!


What order does ISO standard follow? (e.g. year to second, second to year…) - YYYY-MM-DD. I.e. Years to Seconds - Largest to smallest.

What is the minimum number of digits you can have in a year? Why is this? - 4 Digits in the year notation. This is to avoid the "Millenium Bug"

What clock system do the times use? - UTC (Universal Coordinated Time), 24 hour clock

How many time formats does ISO standard have? - hh:mm:ss.sss	or	hhmmss.sss or  hh:mm:ss	or	hhmmss or hh:mm	or	hhmm or hh	

What timezone are time and dates automatically represented in? - UTC 0

How are combined date-times represented? - 


```{r}
library(lubridate)
```
```{r}
class(today())
class(now())
```
```{r}
today()
now()
```
# extract dates out of strings: 
- ymd
- mdy
- dmy
```{r}
"2019-01-19"
ymd("2019-01-19")

```
```{r}
"January 19th, 2019"
mdy("January 19th, 2019")
```
```{r}
"09-Jan-2019"
dmy("09-Jan-2019")
```
Include times: 
- ymd_hms
etc


# Task:
Use the lubridate package to turn the following dates into dates or datetimes:
“March 12th, 2005”
“2018-06-23"
“2017-04-29 11:43:02”
“02-Sep-2002"
“2001-01-07 7”
“August 1st, 2019 08:13"
```{r}
mdy("March 12th, 2005")
ymd("2018-06-23")
ymd_hms("2017-04-29 11:43:02")
dmy("02-Sep-2002")
ymd_h("2001-01-07 7")
mdy_hm("August 1st, 2019 08:13")
```


```{r}
date_components <- tibble(id = 1:5, year = 2013 , month = 3:7, day = 1, hour = 5, minute = id^2 + month)
date_components
```
```{r}
library(dplyr)
```
```{r}
date_components %>%
  mutate(date_combo = make_datetime(year, month, day, hour, minute))
date_components
```
```{r}
install.packages("nycflights13")
library(nycflights13)
```

```{r}
nycflights13::flights
```
```{r}
flights %>%
  mutate(departure = make_datetime(year, month, day, hour, minute))
```
```{r}
as_datetime(today())
as_date(now())
```
```{r}
last_month <- "2019-07-01 04:23:12"
as_datetime(last_month) # or
ymd_hms(last_month)
date(last_month)
```

```{r}
our_datetime <- ymd_hms("2019-06-09 21:04:56")
our_datetime
```
extraction functions, yday starts counting from Sunday.
```{r}
year(our_datetime)
month(our_datetime, label = TRUE, abbr = FALSE)
day(our_datetime)
mday(our_datetime)
yday(our_datetime)
wday(our_datetime, label = TRUE, abbr = FALSE, week_start = 1)
hour(our_datetime)
minute(our_datetime)
second(our_datetime)
```

# timespans

* period:

```{r}
as.period(today() - ymd(19830529))
```
```{r}
today() - years(1)
today() - days(100)
now() - hours(2)
```
```{r}
xsmas <- c("25-12-2020")
xsmas <- dmy(xsmas)
xsmas - today()
xsmas + months(6)
xsmas - weeks(2)
xsmas + hours(36)
xsmas + hours(24)

```

* duration:

as.duration() or dyear or similar

```{r}
ymd(20200110) + years(1) # context dependent (daylight savings times, leap years seconds), i.e 365 days later
ymd(20200110) + dyears(1) # not context dependent - i.e. same day next year
```

* interval: 

```{r}
my_interval <- interval(today(), ymd(20190621))
new_interval <- interval(today(), ymd(20190619))

int_overlaps(my_interval, new_interval)
setdiff(my_interval, new_interval)
```

# timezones

```{r}
Sys.timezone()
OlsonNames()
```

```{r}
ymd_hms("2020-06-02 04:00:00", tz = "Pacific/Auckland")
```

```{r}
with_tz(now(), tzone = "Pacific/Auckland")
```

# task:

Let’s say a plane leaves Sydney Australia at 2.15pm 7-August-2018, and arrives at 5.40am on the 8-August-2018 in London.

Create these two dates with the appropriate time zones.
Work out how long the duration between take off and landing is.

If you left London at 6.20 am on the 10-August-2018, and the total flight time including stop overs was 28 hours and 46 minutes. What time would you arrive in Sydney, Australia?

```{r}
depart_sydney <- ymd_hm("2018-08-07 14:15", tz = "Australia/Sydney")
arrive_london <- ymd_hm("2018-08-08 05:40", tz = "Europe/London")
flight_duration <- arrive_london - depart_sydney
flight_duration <- as.period(flight_duration)

dep_london <- ymd_hm("2018-08-10 06:20", tz = "Europe/London")
duration <- hm("28:46")
arr_sydney <- dep_london + duration
arr_sydney <- with_tz(arr_sydney, tzone = "Australia/Sydney")


flight_duration
arr_sydney
```

# Time series analysis intro.
```{r}
install.packages("tsibble")
install.packages("tsibbledata")
install.packages("feasts")
install.packages("fable")
```
```{r}
library(tsibble)
library(tsibbledata)
library(feasts)
library(fable)
library(tidyverse)
```

```{r}
dim(vic_elec)
```
filter where data was collected in 2012
```{r}
vic_elec %>%
  filter_index("2012")
```
grouping observations
```{r}
vic_year <- vic_elec %>%
  mutate(year = lubridate::year(Date))
vic_year
```
```{r}
vic_year %>%
  index_by(year)%>%
  summarise(mean_temp = mean(Temperature))
```

```{r}
holidays <- vic_elec %>%
  filter(Holiday == FALSE)
```
```{r}
ggplot(holidays) + 
  aes(x = Time, y = Temperature) +
  aezgeom_line()
```

* aggregating

```{r}
# aggregate by date
elec_date <- vic_elec %>%
  index_by(date = ~ as_date(.)) %>%
  summarise(temp_mean = mean(Temperature, na.rm = TRUE))
elec_date
```

```{r}
ggplot(elec_date) + 
  aes(x = date, y = temp_mean) + 
  geom_line()
```
```{r}
elec_month <- vic_elec %>%
  index_by(date = ~ month(., label = TRUE)) %>%
  summarise(temp_mean = mean(Temperature, na.rm = TRUE))
```
```{r}
ggplot(elec_month) +
  aes(x = date, y = temp_mean) +
  geom_line(group = 1) +
  geom_point()
```
* Rolling windows

```{r}
elec_rolling <- vic_elec %>% 
  mutate(temp_ma = slide_dbl(Temperature, 
                             ~ mean(., na.rm = TRUE), 
                             .size = 1000)
         )
```
```{r}
ggplot(elec_rolling) + 
  geom_line(aes(x = Date, y = Temperature), color = "grey") +
  geom_line(aes(x = Date, y = temp_ma),color = "red")
```
```{r}
#feasts
holidays <- tourism %>%
  filter(Purpose %in% c("Holiday","Business")) %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
holidays

```
```{r}
holidays %>%
  autoplot(Trips) + 
  xlab("year")
```
```{r}
holidays %>%
  filter(State %in% c("Queensland", "New South Wales", "Victoria")) %>%
  gg_season(Trips)
```

```{r}
holidays %>%
  filter(State %in% c("Queensland", "New South Wales", "Victoria")) %>%
  gg_subseries(Trips)
```








