---
title: "Inferential Stats, Sampling & distributions"
output: html_notebook
---
# Sampling
```{r}
library(tidyverse)
library(janitor)
```

```{r}
telco <- read_csv("1_sampling_distributions_standard_errors/data/telecomms_churn.csv")
telco <- clean_names(telco)
glimpse(telco)
```
Variables: 

monthly_charges, tenure, churn
```{r}
summary_popn <- telco %>%
summarise(mean_monthly_charges = mean(monthly_charges),
          mean_tenure = mean(tenure),
          prop_churn = mean(churn == "Yes")
          )
summary_popn
```
```{r}
telco %>%
  ggplot(aes(x = monthly_charges)) +
  geom_histogram(col = "white")
```
monthly charges strongly bimodal
```{r}
telco %>%
  ggplot(aes(x = tenure)) +
  geom_histogram(col = "white")
```
not a normal distribution
```{r}
telco %>% 
  ggplot(aes(x = churn)) +
  geom_bar()
```

ABOVE EVERYTHING IS POPULATION DISTRIBUTIONS. 


# take random sample of 200 records from telco dataset.

SRS = simple random sample
install "infer" package

```{r}
library(infer)
```
```{r}
sample_200 <- telco %>%
  rep_sample_n(size = 200, reps = 1)
sample_200
```

Below are estimates of the telco population aka statistic or point estimates.

```{r}
summary_sample_200 <- sample_200 %>%
  ungroup() %>%
  summarise(mean_monthly_charges = mean(monthly_charges),
          mean_tenure = mean(tenure),
          prop_churn = mean(churn == "Yes")
          )
summary_sample_200
  
```

Sampling error = difference between parameters and statistics

```{r}
summary_sample_200 - summary_popn
```

Repeat sampling = Resampling
5000 resamples each or size 200.

Below are sampling distributions of statistics.
```{r}
rep_sample_200 <- telco %>%
  rep_sample_n(size = 200, reps = 5000) %>%
  summarise(mean_monthly_charges = mean(monthly_charges),
          mean_tenure = mean(tenure),
          prop_churn = mean(churn == "Yes")
  )
rep_sample_200
```

Mean monthly charges sampling distribution

```{r}
mean_monthly_charges_plot <- rep_sample_200 %>%
  ggplot(aes(x = mean_monthly_charges)) + 
  geom_histogram(col = "white") + 
  labs(x = "mean monthly charges from each sample")
mean_monthly_charges_plot
```

```{r}
tenure_sample200_plot <- rep_sample_200 %>%
  ggplot(aes(x = mean_tenure)) + 
  geom_histogram(col = "white") + 
  labs(x = "mean tenure from each sample")
tenure_sample200_plot
```
```{r}
tenure_sample200_plot <- rep_sample_200 %>%
  ggplot(aes(x = prop_churn)) + 
  geom_histogram(col = "white") + 
  labs(x = "Proportion of churn from each sample")
tenure_sample200_plot
```

Central limit theorem: 

The larger the amount of data and the more samples, the more normal is the distribution.

Proportions need to satisfy success/failure criteria - 

p = proportion of "successes" 1-p = proportion of "failures". 

n * p >= 10 (number of success)
n * (1-p) >= (number of failures)

p ~ 0.01, what sample size is needed?

n * 0.01 >= 10


Rule of thumb - a sample of 30 is usually good enough.

standard error = standard deviation of sample distribution. 

```{r}
#standard errors
```


```{r}
std_errs <- rep_sample_200 %>%
  summarise(
    se_mean_monthly_charges = sd(mean_monthly_charges),
    se_mean_tenure = sd(mean_tenure),
    se_prop_churn = sd(prop_churn),
  )
std_errs
```
```{r}
rep_sample_200 %>%
  filter(mean_monthly_charges >= 60, mean_monthly_charges <= 70) %>%
  summarise(prob = n() / nrow(rep_sample_200))
```


# BOOTSTRAPPING.
```{r}
library(janitor)
library(tidyverse)
```
```{r}
sample_200 <- read_csv("5_confidence_intervals/data/telecomms_churn_200_sample.csv")
```

```{r}
bootstrap_resample_200 <- sample_200 %>%
  rep_sample_n(size = 200, replace = TRUE, reps = 5000) %>%
  summarise(mean_tenure = mean(tenure))
bootstrap_resample_200

  ggplot(bootstrap_resample_200) +
    aes(x = mean_tenure) + 
  geom_histogram(col = "white")


sd(bootstrap_resample_200$mean_tenure)
```

# confidence interval
```{r}
ci_95 <- bootstrap_resample_200 %>%
  summarise(mean = mean(mean_tenure),
            lower_bound = quantile(mean_tenure, probs = 0.025),
            upper_bound = quantile(mean_tenure, probs = 0.975)
            )
ci_95
```


### INFER : 
- Specify,
- Generate,
- Calculate

```{r}
infer_resample <- sample_200 %>%
  specify(response = tenure) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

infer_resample
```
```{r}
infer_ci_95 <- infer_resample %>%
  get_ci(level = 0.95, type = "percentile")
infer_ci_95
```
```{r}
infer_resample %>%
  visualise() +
  shade_confidence_interval(endpoints = infer_ci_95)
```
```{r}
infer_resample_300 <- read_csv("5_confidence_intervals/data/telecomms_churn_300_sample.csv")
```

```{r}
infer_resample_300 %>%
  specify(response = tenure) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean") %>%
  get_ci(level = 0.95, type = "percentile")

infer_resample_300

infer_resample_400 <- read_csv("5_confidence_intervals/data/telecomms_churn_400_sample.csv") %>%
  specify(response = tenure) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean") %>%
  get_ci(level = 0.95, type = "percentile")

infer_resample_400

infer_resample_500 <- read_csv("5_confidence_intervals/data/telecomms_churn_500_sample.csv") %>%
  specify(response = tenure) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean") %>%
  get_ci(level = 0.95, type = "percentile")

infer_resample_500
```


```{r}

```



