---
title: "Untitled"
author: "Gleb V"
date: "20/07/2020"
output: html_document
---
yaml: yet another markup language
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Markdown is a very simple markup language
<h1>This would be the largest heading</h1>
# This is the markdown equivalent
<h6>Smaller heading</h6>
###### Equivalent in markdown
\# escapes
# Emphasis
<b>HTML bold</b>
**Markdown bold**  
__Equivalent__
<i>Italics in HTML</i>
*Markdown italics*
_Equivalent_
# Lists
<ol>
  <li>These elements</li>
  <li>... will be ...</li>
  <li>numbered</li>
</ol>
<ul>
  <li>These won't</li>
  <li>be numbered</li>
</ul>
1. The first item
2. The second item
    * Notice this
    * is indented
3. Back to numbering
# Links
<a href="www.google.com">HTML links are annoying</a>
[Markdown links are a little bit better](www.google.com)
* Bullet
* points
# Task
# Welcome to our page!
  
This is an example of how we **could** lay something out.
  
### Smaller headings are fun too!
* This element is part of a list
* So is this one
* [This one's a link too!](https://http.cat/200)
# Maths
LaTex language for writing maths $\frac{2n}{n+1}$
$$2x^2+x+4$$
# Code
```{r}
x <- 4
```
```{r echo=FALSE}
# statistical analysis
y <- mean(rnorm(100))
```
The result was `r y`, neat huh?
# Tables
```{r}
library(knitr)
kable(mtcars)
```
| Name | Species | Age |
|---|---|---|
| Ella  | Tortoise | 6 |
| Poppy | Dog | 2 |
| Millie | Cat | 10 | 
