---
title: "R_Markdown_intro"
author: "Gleb V"
date: "20/07/2020"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is how to make a sidebar menu

`output: 
  html_document:
    toc: true
    toc_float: true`

language used in R markdown is "Markdown": https://www.markdownguide.org/getting-started/

<h1>This is lv1 HTML heading</h>

#\# hashtag used to define headings in Markdown h1 = \#, h6 = \#\#\#\#\#\#\#


# Emphasis 

Bold = 

**Markdown bold**

__Is the same__

italics = 

*italics*

_italics_

# Lists

## Ordered list 

1. This is

2. a Markdown 

2.1 ordered list

  * This is 
  * unordered list
  
# Links

[Markdown link text](www.google.com)

# TASK

# Welcome to our page!

This is an example of how we __could__ lay something out. 

### Smaller headings are fun too!

* This element is part of a list
* So is this one
* [This one is a link too!](https://http.cat/200)


# WRITING MATHS WITH LaTex

$\frac{2n}{n+1}$
$$2xˆ2+x+4$$
# REFERENCING A VARIABLE
```{r echo=FALSE}
y <- 3
```
The value of backtick y backtick should render the value stored in the variable y



# Tables
```{r}
mtcars
```

## Manual tables

| Name | Species | Age | 
| --- | ---| --- |
| Ella | Tortoise | 6 | 
| Poppy | Dog | 2 | 
| Millie | Cat | 10 | 

Table gerator website:[https://www.tablesgenerator.com/markdown_tables](https://www.tablesgenerator.com/markdown_tables)
