---
title: "R Notebook"
output: html_notebook
---
```{r}
library(RPostgreSQL)
```

```{r}
home_location <- Sys.getenv("HOME")
credentials_location <- paste0(home_location,"/credentials.R")
source(credentials_location)
```

```{r}
db_connect <- dbConnect(drv = PostgreSQL(), user = postgresql_username, password = postgresql_password, dbname = "omni_employees", host = "localhost")
rm(postgresql_username, postgresql_password)
```
SQL Database querying in R
```{r}
#get full dataframe
dbGetQuery(conn = db_connect, statement = 
             "SELECT * FROM employees;")
```
```{sql connection=db_connect}
--SQL comment
SELECT * FROM employees;
```
select team_id 1
```{sql connection=db_connect}
SELECT * FROM employees WHERE team_id = 1;
```
working 0.5 fre or more
```{sql connection=db_connect}
SELECT * FROM employees WHERE fte_hours >= 0.5;
```
all employees not based in Brazil
```{sql connection=db_connect}
SELECT * FROM employees WHERE country != 'Brazil';
```
find all employees in China that started working in 2019
```{sql connection=db_connect}
SELECT * FROM employees
WHERE country = 'China' AND start_date >= '2019-01-01';
```

find all employees in China that started working after 2019 or are enrolled in pension
```{sql connection=db_connect}
SELECT * FROM employees
WHERE country = 'China' AND (start_date >= '2019-01-01' OR pension_enrol = TRUE);
```
find all employees working between 0.25 and 0.5 fte hours inclusive
```{sql connection=db_connect}
SELECT * FROM employees
WHERE fte_hours >= 0.25 AND fte_hours <= 0.5;
```

find employees that started working in years other than 2017

```{sql connection=db_connect}
SELECT * FROM employees
WHERE start_date <= '2017-01-01' OR start_date >= '2017-12-31';
```

find all employees working between 0.25 and 0.5 fte hours inclusive USING BETWEEN.
```{sql connection=db_connect}
SELECT * FROM employees
WHERE fte_hours BETWEEN 0.25 AND 0.5;
```

find employees that started working in years other than 2017 using BETWEEN

```{sql connection=db_connect}
SELECT * FROM employees
WHERE start_date NOT BETWEEN '2017-01-01' AND '2017-12-31';
```

find employees that started working in 2016 and work 0.5 fte or more
```{sql connection=db_connect}
SELECT * FROM employees
WHERE (start_date BETWEEN '2016-01-01' AND '2016-12-31') AND (fte_hours >= 0.5);
```

find all employees in Spain, South Africa, Ireland and Germany.
```{sql connection=db_connect}
SELECT * FROM employees
WHERE country IN('Spain', 'South Africa', 'Ireland', 'Germany');
```

NOT IN Finland Argentina Canada
```{sql connection=db_connect}
SELECT * FROM employees
WHERE country NOT IN('Finland', 'Argentina', 'Canada');
```
find all whos last name begins with 'Mc..'
```{sql connection=db_connect}
SELECT * FROM employees
WHERE country = 'Greece' AND last_name LIKE 'Mc%'
-- % is a wildcard

```
find all with no email
```{sql connection=db_connect}
SELECT * FROM employees
WHERE email IS NULL
```

selecting specific columns only
```{sql connection=db_connect}
SELECT id, first_name, last_name FROM employees
WHERE department = 'Accounting'
```

first and last names combined in one column
```{sql connection=db_connect}
SELECT 
  id,
  first_name,
  last_name,
  CONCAT(first_name, ' ', last_name) AS full_name
FROM employees
```

filter rows that don't have both first and last name
```{sql connection=db_connect}
SELECT 
  id,
  first_name,
  last_name,
  CONCAT(first_name, ' ', last_name) AS full_name 
FROM employees
WHERE first_name IS NULL OR last_name IS NULL;
```

# DISTINCT values

```{sql connection=db_connect}
SELECT 
DISTINCT(department)
FROM employees
```

# AGGREGATE functions

How many STARTED  WORKING in 2001
```{sql connection=db_connect}
SELECT
  COUNT(*) AS total_started_in_2001
FROM employees
WHERE EXTRACT(YEAR FROM start_date) = 2001
```

Maximum and minimum salaries of all employees?
```{sql connection=db_connect}
SELECT
  MIN(salary) AS min_salary,
  MAX(salary) AS max_salary
FROM employees
```

Average salary of employees in the Human Resources.
```{sql connection=db_connect}
SELECT
  AVG(salary) AS avg_HR_salary
FROM employees
WHERE department = 'Human Resources'

```

How much does the corporation spend on the salaries of employees hired in 2018
```{sql connection=db_connect}
SELECT 
  SUM(salary) AS total_spend_on_salaries
FROM employees
WHERE EXTRACT(YEAR FROM start_date) = 2018

```

# Order. 

```{sql connection=db_connect}
SELECT *
FROM employees
WHERE salary IS NOT NULL
ORDER BY salary DESC
LIMIT 1
```

```{sql connection=db_connect}
```


```{sql connection=db_connect}
SELECT * 
FROM employees
ORDER BY salary ASC NULLS LAST 
LIMIT 1
```

longest serving employee
```{sql connection=db_connect}
SELECT *
FROM employees
ORDER BY start_date ASC NULLS LAST 
LIMIT 1
```

Highest paid employee in Libya
```{sql connection=db_connect}
SELECT * 
FROM employees
WHERE country = 'Libya'
ORDER BY salary DESC NULLS LAST
LIMIT 1
```
most commom fte across the corp, based in America.
```{sql connection=db_connect}
SELECT 
  fte_hours,
  COUNT(id) AS total
FROM employees 
WHERE fte_hours IS NOT NULL AND country = 'United States'
GROUP BY fte_hours 
ORDER BY COUNT(fte_hours) DESC
-- LIMIT 1
```

NUMBER of employees within each department
```{sql connection=db_connect}
SELECT
  department,
  COUNT(id) AS number_of_employees
FROM employees
GROUP BY department
```

how many employees in each country
```{sql connection=db_connect}
SELECT 
  country,
  COUNT(id) AS num_employees
FROM employees
GROUP BY country
```
how many in each department work either 0.25 or 0.5 fte

```{sql connection=db_connect}
SELECT 
  department,
  COUNT(id) AS num_of_employees
FROM employees
WHERE fte_hours BETWEEN 0.25 AND 0.5 -- could do IN(0.25, 0.5) instead of BETWEEN
GROUP BY department
```

how many employees in each dept are enrolled in pension
```{sql connection=db_connect}
SELECT
  department,
  COUNT(id) AS num_employees
FROM employees
WHERE pension_enrol IS TRUE
GROUP BY department
```
breakdown employees by country where first name is NULL
```{sql connection=db_connect}
SELECT 
  country,
  COUNT(id) AS num_employees
FROM employees
WHERE first_name IS NULL
GROUP BY country
```

depts where at lest 40 employees work either 0.25 or 0.5 fte

```{sql connection=db_connect}
SELECT 
  department,
  COUNT(id) AS num_of_employees
FROM employees
WHERE fte_hours IN(0.25, 0.5)
GROUP BY department
HAVING COUNT(id) >= 40

```
```{sql connection=db_connect}
SELECT
  country,
  MIN(salary) ASmin_salary
FROM employees
WHERE pension_enrol IS TRUE
GROUP BY country
HAVING MIN(salary) < 21000
```
# SUBQUERIES

Japan employees earning over company wide average salary

```{sql connection=db_connect}
SELECT *
FROM employees
WHERE country = 'Japan' AND salary > (SELECT AVG(salary) FROM employees)
```

depts where at lest 40 employees work either 0.25 or 0.5 fte as subquery
```{sql connection=db_connect}
SELECT *
FROM employees
WHERE country = 'United States' AND fte_hours =
  (SELECT fte_hours
   FROM employees
   WHERE fte_hours IS NOT NULL
   GROUP BY fte_hours
   ORDER BY COUNT(fte_hours) DESC
   LIMIT 1
  )

```


```{r}
#disconnect from database in the end
dbDisconnect(db_connect)
```


