---
title: "R Notebook"
output: html_notebook
---

```{r}
numbers <- c(1, 4, 6, 3)
```
```{r}
numbers
```
```{r}
more_numbers <- c(1,1,6,2,3,3,2,6)
many_numbers <- c(numbers, more_numbers)
many_numbers
```
```{r}
primes <- c(2,3,5,7,11)
primes <- c(primes,13,17)
```

```{r}
class(numbers)
```
```{r}
class(c(1L, 5L, 6L)) #letter L specifies each element as Integer, by default numeric elements are "doubles"
```
```{r}
class(c("a", "a", "b", "a")) #character data
```
```{r}
as.factor(c("a", "a", "b", "a")) #saves as factor, factors are ordered
class(as.factor(c("a", "a", "b", "a")))
```
```{r}
letters <- factor(c("a", "a", "b", "a"), ordered = TRUE)
letters[3] > letters[2]
```
```{r}
class(c(TRUE, FALSE, FALSE, TRUE))
```
```{r}
tolower(LETTERS)
```
```{r}
numbers + 5
```
```{r}
c(1, 2) + 3
#is the same as 
c(1, 2) + c(3, 3)
```
```{r}
as.character(numbers)
as.numeric(c("1", "2"))
as.logical(c(0, 1, -10))
```
```{r}
sum(numbers)
```
```{r}
mean(numbers)
```
```{r}
length(numbers)
```
```{r}
rev(numbers) #reverse order
```
```{r}
many_numbers
unique(many_numbers) # returns only unique elements of the verctor
table(many_numbers) # tells how many times each element appears in a verctor
sort(many_numbers) # arranges elements in order, for characters it is alphabetical order
rev(sort(LETTERS))
```
```{r}
1:5
```
```{r}
seq(0, 50, by = 2) # creates a specifiyq sequence with a specified step
```
```{r}
rep(c(1, 2, 3), times = 4) #repeats a vector specified number of times
rep(c(1, 2, 3), each = 4) #repeats each element
```

Create a sequence going from 10 to 100 in steps of 5. Name this vector answer_1.

Make a vector which repeats the number 8 six times. Name this vector answer_2.

Use rep() to make a vector like below. Name the vector answer_3.
## [1] 2 2 2 3 3 3 4 4 4

Make a vector by joining the answer_1, answer_2 and answer_3 vectors together.

What does sort() do to a logical vector?

Compare the sum of answer_1 with the sum of answer_1 reversed.

What happens when you call table() on a vector of unique elements.

What does sum() do to a logical vector?

```{r}
answer_1 <- seq(10, 100, by = 5)
answer_2 <- rep(8, times = 6)
answer_3 <- rep(c(2, 3, 4), each = 3)
vec <- c(answer_1, answer_2, answer_3)
sort(as.logical(vec)) # NEEDS AN ACTUAL LOGICAL VECTOR
sum(rev(answer_1)) == sum(answer_1)
table(unique(answer_3))
```
```{r}
sum(as.logical(vec))
```
Arithmetic operators

Brackets
Of (powers)
Ddivide
Multiply
Add
Subtract
```{r}
5 + 5
3^2
6 - 1
2 * 5
10 / 2

#integer division
10 %% 3
#operators work on vectors
1:20 + 5
```

```{r}
6 == 6
c(6, 5, 6, 7) == 6
```
```{r}
6 != 6
```
```{r}
c(6, 5, 6, 7) < 6
c(6, 5, 6, 7) <= 6
c(6, 5, 6, 7) >= 6
```
```{r}
5 %in% 1:6 #checks if a number appears in a vector
```
```{r}
!c(TRUE, FALSE, TRUE)
```
```{r}
TRUE & TRUE
FALSE & FALSE
TRUE | FALSE
FALSE | FALSE
```
Locical operators written like & and | obey vector recycling, && and || do not use vector recycling. 

```{r}
names(numbers) <- c("first", "second", "third", "fourth")
numbers
numbers[1:3]
numbers[-1:-3]
```
```{r}
greater_than_three <- numbers > 3
numbers[greater_than_three]
```
Select the third element of answer_1.

Select everything except the last element of answer_1.

Select all the elements of answer_1 which are less than 25.
```{r}
answer_1[3]
answer_1[-length(answer_1)]
answer_1[answer_1 < 25]
```

























