---
title: "Distributions"
output: html_notebook
---

```{r}
library(prob)
library(tidyverse)
library(janitor)
library(stats)
```

```{r}
s_three_coins <- tosscoin(3)
s_three_coins
```
```{r}
s_three_coins <- s_three_coins %>%
  mutate(label = paste(toss1, toss2, toss3, sep = "")) %>%
  mutate(prob = 1 / n())
s_three_coins
```
```{r}
ggplot(s_three_coins) + 
  aes(x = label, y = prob) + 
  geom_col()
```
This is an even distribution

```{r}
s_three_coins <- s_three_coins %>%
  mutate(num_heads = rowSums(. == "H"))
s_three_coins
```
```{r}
prob_heads <- s_three_coins %>%
  group_by(num_heads) %>%
  summarise(prob = sum(prob))
prob_heads
```
```{r}
ggplot(prob_heads) + 
  aes(x = num_heads, y= prob) +
  geom_col()
```
#
#
# Bunch of missing code about coin flipping and frequency distro plot
#
#
#
```{r}
air_con_sales <- read_csv("1_distributions_measures_of_centre/data/AirConSales.csv")
```
```{r}
air_con_sales$Date <- as.Date(air_con_sales$Date, format = "%m/%d/%Y")
air_con_sales
```
```{r}
sales_freq_table <- table(air_con_sales$Units_sold)
sales_freq_table
```
```{r}
rolls <- sample(1:6, 100, replace = TRUE)
freq <- table(rolls)
freq

rel_freq <- freq /100
rel_freq
```
```{r}
daily_sales_dist <- air_con_sales %>%
  tabyl(Units_sold)
daily_sales_dist
```
```{r}
ggplot(daily_sales_dist) +
  aes(x = Units_sold) + 
  geom_histogram(bins = 20, col = "white")
```

# Mean

```{r}
air_con_sales %>%
  summarise(mean_daily_sales = mean(Units_sold))
```

# Median

```{r}
air_con_sales %>%
  summarise(mean_daily_sales = median(Units_sold))
```

# Mode

```{r}
# Mode function

getMode <- function(data){
  tabled_data <- table(data)
  table_names <- names(tabled_data)
  return( table_names[tabled_data == max(tabled_data)] )
}
```
```{r}
air_con_sales %>%
  summarise(mode_daily_sales = getMode(Units_sold))
```
```{r}
air_con_sales <- boxplot(air_con_sales$Units_sold)
```

# Unimodal & bimodal distros - distributions with different number of peaks.

```{r}
unimodal <- read_csv("1_distributions_measures_of_centre/data/unimodal.csv")
```
unimodal distribution
```{r}
ggplot(unimodal) +
  aes(x = x) +
  geom_histogram(bins = 20, col = "white")
```
bimodal dist
```{r}
bimodal <- read_csv("1_distributions_measures_of_centre/data/bimodal.csv")
```
```{r}
ggplot(bimodal) + 
  aes(x = x) +
  geom_histogram(bins = 20, col = "white")
```

# Skewedness

How assymetric is data.

```{r}
left_skew <- read_csv("1_distributions_measures_of_centre/data/leftskew.csv")
ggplot(left_skew) +
  aes(x = x) +
  geom_histogram(bins = 20, col = "white")
```
```{r}
right_skew <- read_csv("1_distributions_measures_of_centre/data/rightskew.csv")
ggplot(right_skew) +
  aes(x = x) +
  geom_histogram(bins = 20, col = "white")
```
Skew is where the tail is - where the least data pointing to. 

# 2. Distributions measures of spread

```{r}
jobs <- read_csv("2_distributions_measures_of_spread/data/TyrellCorpJobs.csv")

```
```{r}
summary(jobs)
```
## Range max value - min values
```{r}
jobs %>%
  group_by(Position) %>%
  summarise(min = range(Salary)[1],
            max = range(Salary)[2],
            range = max - min)

```
```{r}
ggplot(jobs) +
  aes(x = Salary) +
  geom_histogram(bins = 25, col = "white") + 
  facet_wrap(~ Position)
```

# QRs & IQR

IQR = Q3 - Q1
Q2 = median

```{r}
jobs %>%
  group_by(Position) %>%
  summarise(Q1 = quantile(Salary, 0.25),
            Q2 = quantile(Salary, 0.55),
            Q3 = quantile(Salary, 0.75),
            IQR = Q3 - Q1
            )
```
```{r}
install.packages("skimr")
library(skimr)
```
```{r}
jobs %>%
  select(-X1) %>%
  group_by(Position) %>%
  skim()
```

# Variance and Standard Deviation.

```{r}
jobs %>%
  group_by(Position) %>%
  summarise(variance = var(Salary)) #variance values are squared
```
```{r}
jobs %>%
  group_by(Position) %>%
  summarise(stdev = sd(Salary))
```


# Probability distributions - continuous is different from descreet. 

## Descreet

## Continuous

```{r}
punif(10.751, min = 5, max = 22) - punif(8.4, min = 5, max = 22) # what is punif?


dunif
```

```{r}
runif(100, 5, 22) # sample of uniform distribution, for continuous data only.
#can use smple() for descreet values
```

# !!!!!!!!!!!!!!!!Normal distribution!!!!!!!!!

```{r}

```

