---
title: "GG Plot intro"
output: html_notebook
---
```{r}
library(ggplot2)
library(CodeClanData)
```
```{r}
students
```
## basics.
```{r}
ggplot(students) + 
  geom_bar(aes(x = superpower, fill = school_year), position = "fill") + 
  scale_fill_brewer()
```
```{r}
ggplot(students) + 
  geom_bar(aes(x = superpower, fill = school_year), position = "dodge") + 
  scale_fill_brewer()
```

## statistical transformation
```{r}
ggplot(students) + 
  geom_bar(aes(x = superpower, fill = school_year), stat = "count") + 
  scale_fill_brewer()
```
```{r}
library(tidyverse)

count_data <- students %>%
  group_by(superpower, school_year) %>%
  summarise(counts = n())

count_data %>%
  ggplot() + 
  geom_bar(aes(x = superpower, y = counts, fill = school_year), stat = "identity")
#stat = "identity" to use if I want to keep a variable (in this case y = counts) the same.
```

geom_bar - used if data hasn't been precounted
geom_col - if data has been precounted

## labels

```{r}
ggplot(students) + 
  geom_bar(aes(x = superpower, fill = school_year)) + 
  labs(x = "Superpower", 
       y = "Count", 
       title = "Preferred superpower by school year", 
       subtitle = "data from students around the world", 
       fill = "school year") + 
scale_fill_brewer()
```
```{r}
ggplot(students) +
  geom_point(aes(x = height_cm, y = reaction_time, colour = superpower)) + 
  geom_jitter(aes(x = height_cm, y = reaction_time, colour = superpower), width = 5, height = 0.5) 
```


```{r}
top_10 <- olympics_overall_medals %>%
  arrange(count) %>% 
  top_n(10)
top_10

ggplot(top_10) + 
  geom_bar(aes(x = team, y = count, fill = medal), stat = "identity") + 
  labs(x = "Team", 
       y = "Count of medals", 
       title = "Count of Olympic medals by team",
       fill = "Medals") + 
scale_fill_brewer()
```
This plot summarises the number of medals each team had. On X axis are the team names, Y axis shows total number of each type of medals each team had 

```{r}
data("ChickWeight")
head(ChickWeight)
```
Track the weight of each chick in time, color lines by die time.

```{r}
ggplot(ChickWeight) + 
  #a common aes() can be added to avoid redundancy.
  #aes(x = Time, y = weight, colour = Diet)
  geom_line(aes(x = Time, y = weight, group = Chick, colour = Diet), alpha = 0.1
            ) + 
  geom_point(aes(x = Time, y = weight, colour = Diet), alpha = 0.2
             ) + 
  geom_smooth(aes(x = Time, y = weight, colour = Diet))
```
```{r}
head(pets)

ggplot(pets) + 
  aes(x = age, y = weight) + 
  geom_point(aes(colour = sex), size = 5) +
  geom_text(aes(label = name), 
            nudge_x = 0.5, 
            nudge_y = 0.1)

ggsave("g1_sav.pdf")
```


# Facets
One plot converted into many plots using facet_wrap() or facet_grid()

```{r}
ggplot(ChickWeight) + 
  aes(x = Time, y = weight, group = Chick, colour = Diet) + 
  geom_line() + 
  facet_wrap(~ Diet, ncol = 4) # ~ means "varis with" or "depending on"
```
```{r}
benefits

ggplot(benefits) +
  aes(age) +
  geom_histogram(col = "white")  + 
  facet_wrap(~sex, nrow = 2)
```

```{r}
ggplot(benefits) +
  geom_histogram(aes(x = age, fill = sex), col = "white") + 
  facet_grid(married ~ sex) #rows ~ columns
```
```{r}
ggplot(benefits) + 
  geom_histogram(aes(x = age), col = "white", fill = "dark green") + 
  facet_grid(married ~ dkids, labeller = "label_both")
```

```{r}
ggplot(benefits) + 
  geom_histogram(aes(x = age), col = "white", fill = "dark green") + 
  facet_wrap(~sex, ncol = 2, scales = "free")
```
## bins / binning

cut_iterval(x, n) - divides data into n bins, each bin same width.
cut_width(x, width) - divides data into n bins, each bin of width wide.
cut_number(x, n) - divides data into n n bins. 
```{r}
library(tidyverse)

```
```{r}
benefits <- benefits %>%
  mutate(
    tenure_interval = cut_interval(tenure, 4), 
    tenure_width = cut_width(tenure, 15),
    tenure_number = cut_number(tenure, 4)
  )

ggplot(benefits) + 
  geom_histogram(aes(x = age), col = "white") + 
  facet_wrap(~tenure_interval)

ggplot(benefits) + 
  geom_histogram(aes(x = age), col = "white") + 
  facet_wrap(~tenure_number)
```
# Aesthetics

```{r}
ggplot(pets) + 
  aes(age, weight, colour = sex) + 
  geom_point() + 
  scale_x_continuous(breaks = c(1, 6, 10), 
                     labels = c("1" = "Baby",
                                "6" = "Grown up",
                                "10" = "Over the hill")) + 
  scale_y_continuous(breaks = c(6,8,10),
                     labels = c("6" = "Small",
                                "8" = "Medium",
                                "10" = "Large"))
```
```{r}
library(scales)
```
```{r}
ggplot(flatPrices) + 
  aes(x = Date, y = AvgFlatPrice) + 
  geom_line() + 
  ylab("average flat price \n") + 
  scale_y_continuous(labels = scales::dollar_format(prefix = "£"))
```


```{r}
flatPrices <- flatPrices %>%
  mutate(percent_change = AvgFlatPrice/AvgFlatPrice[1])

ggplot(flatPrices) + 
  aes(x = Date, y = percent_change) + 
  geom_line() + 
  scale_y_continuous(labels = scales::percent) + 
  ylab("percent change relative to start \n")
```

### Changing limit of scales.

```{r}
ggplot(pets, aes(age, weight)) +
  geom_point(aes(colour = sex), size = 4) +
  scale_x_continuous(
    name = "Age",
    breaks = seq(13),
    limits = c(1,8)
  ) + 
  scale_y_continuous(
    name = "Weight",
    breaks = c(6, 8, 10),
    labels = c(
      "6" = "Light",
      "8" = "Normal",
      "10" = "Heavy"
    ), 
    limits = c(1,8)
  )
```

### Changing scale of the values

```{r}
ggplot(whisky) + 
  aes(x = YearFound, y = Capacity, color = Region) + 
  geom_point() +
  #scale_x_reverse()
  scale_y_continuous(trans = "log10")
```

```{r}
ggplot(whisky) + 
  aes(x = YearFound, fill = Region) + 
  geom_histogram(bins = 30) +
  #coord_flip()
  coord_fixed()
```
### PieCharts

```{r}
snacks <- data.frame(
  group = c("Cakes", "Biscuits", "Jaffa Cakes"),
  count = c(30, 50, 20)
)
ggplot(snacks) + 
  aes(x = " ", y = count, fill = group) +
  geom_bar(width = 0.5, stat = "identity") + 
  xlab("biscuits") + 
  # to make a pie chart : 
  coord_polar("y")
```

### Zoom in to an area

```{r}
ggplot(whisky) + 
  aes(x = YearFound, y = Capacity, colour = Region) + 
  geom_point(aes(colour = Region)) +  
  geom_smooth() + 
  xlim(1940,2000)  
  #coord_cartesian()
```










