---
title: "Hypothesis testing intro"
output: html_notebook
---
```{r}
library(infer)
library(janitor)
library(tidyverse)
```

# One sample mean

```{r}
books <- read_csv("1_one_sample_tests_mean/data/books.csv")
glimpse(books)
```
```{r}
books_tidy <- books %>%
  clean_names() %>%
  filter(!is.na(average_rating)) %>%
  rename(num_pages = number_num_pages)
glimpse(books_tidy)
```
```{r}
books_tidy %>%
  ggplot(aes(x = average_rating)) + 
  geom_histogram(col = "white", bins = 30)

books_tidy %>%
  ggplot(aes(x = average_rating)) +
  geom_boxplot()
```
How does the mean of this sample (sample (not full) of 2020) compare to another year.

```{r}
observed_stat <- books_tidy %>%
  summarise(mean_rating = mean(average_rating))
observed_stat
```

#### Calculating null hypothesis distribution:

```{r}
null_distribution <- books_tidy %>%
  specify(response = average_rating) %>%
  hypothesise(null = "point", mu = 3.93) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")
null_distribution
```
```{r}
null_distribution %>%
  visualise() + 
  shade_p_value(obs_stat = observed_stat$mean_rating, direction = "both")
```
```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat$mean_rating, direction = "both")
p_value
```


# TASK
```{r}
spa_books <- books_tidy %>%
  filter(language_code == "spa")
spa_books
```
```{r}
spa_books_obs_stat <- spa_books %>%
  summarise(mean_rating = mean(average_rating))
spa_books_obs_stat
```



```{r}
spa_books_null_distribution <- spa_books %>%
  specify(response = average_rating) %>%
  hypothesise(null = "point", mu = 3.96) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")
spa_books_null_distribution
```
```{r}
spa_books_null_distribution %>%
  visualise() + 
  shade_p_value(obs_stat = spa_books_obs_stat$mean_rating, direction = "left")
```
```{r}
p_value <- spa_books_null_distribution %>%
  get_p_value(obs_stat = spa_books_obs_stat$mean_rating, direction = "left")
p_value
```

# One sample proportions.

Proportion of books without text reviews.

```{r}
books_tidy %>% 
  group_by(text_reviews_count) %>%
  summarise(prop = n()/nrow(books_tidy)) %>%
  filter(text_reviews_count == 0)
```
```{r}
books_tidy_prop <- books_tidy %>%
  mutate(text_reviews_flag = ifelse(text_reviews_count > 0, "text_review", "no_text_review"))

head(books_tidy_prop)
```
Because we need a proportion, we cannot do bootstrapping, but need simulation.

```{r}
null_distribution <- books_tidy_prop %>%
  specify(response = text_reviews_flag, success = "no_text_review") %>%
  hypothesise(null = "point", p = 0.07) %>%
  generate(reps = 5000, type = "simulate") %>%
  calculate(stat = "prop")
head(null_distribution)
```

```{r}
null_distribution %>%
  visualise()
```
```{r}
observed_stat <- books_tidy_prop %>%
  specify(response = text_reviews_flag, success = "no_text_review") %>%
  calculate(stat = "prop")
observed_stat
```

```{r}
null_distribution %>%
  visualise() + 
  shade_p_value(obs_stat = observed_stat, direction = "both")
```
```{r}
null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "both")
```

H0 = 0.015
Ha < 0.015
sig level is 1%

```{r}
#unique(books_tidy$language_code)
books_tidy %>% 
  group_by(language_code) %>%
  summarise(prop = n()/nrow(books_tidy)) %>%
  filter(language_code == "fre")
```

```{r}
books_tidy_prop_fre <- books_tidy %>%
  mutate(language_flag = ifelse(language_code == "fre", "fre_language", "other_language"))
head(books_tidy_prop_fre)
```
```{r}
fre_null_distribution <- books_tidy_prop_fre %>%
  specify(response = language_flag, success = "fre_language") %>%
  hypothesise(null = "point", p = 0.015) %>%
  generate(reps = 2000, type = "simulate") %>%
  calculate(stat = "prop")

head(fre_null_distribution)
```
```{r}
fre_null_distribution %>%
  visualise()
```
```{r}
fre_observed_stat <- books_tidy_prop_fre %>%
  specify(response = language_flag, success = "fre_language") %>%
  calculate(stat = "prop")
fre_observed_stat
```
```{r}
fre_null_distribution %>%
  visualise() + 
  shade_p_value(obs_stat = fre_observed_stat, direction = "less")
```

```{r}
fre_null_distribution %>%
get_p_value(obs_stat = fre_observed_stat, direction = "less")
```

# Two sample tests 

## Independent samples.

We have two samples of prices (in £s) for 2-bedroom, 14-day holiday apartment lets in Nice and the Algarve. We would like to know whether the mean price of lets in the Algarve is significantly higher than the mean price of comparable lets in Nice?’
alpha = 0.05
H0: mu_algarve - mu_nice = 0
Ha: mu_algarve - mu_nice > 0

```{r}
nice <- read_csv("3_two_sample_tests/data/nice.csv")
algarve <- read_csv("3_two_sample_tests/data/algarve.csv")
```

```{r}
apart_prices <- bind_rows(nice, algarve, .id = "location") %>%
  mutate(location = ifelse(location == "1", "nice", "algarve")) %>%
  select(-X1)

apart_prices
```
```{r}
apart_prices %>%
  ggplot(aes(x = location, y = price)) + 
  geom_boxplot()
```

Null by permutation

```{r}
null_distribution <- apart_prices %>%
  specify(response = price, explanatory = location) %>% #(price ~ location)
  hypothesise(null = "independence") %>%
  generate(reps = 5000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("algarve", "nice"))
null_distribution
```

```{r}
#observed stat
observed_stat <- apart_prices %>%
  specify(price ~ location) %>%
    calculate(stat = "diff in means", order = c("algarve", "nice"))
observed_stat
```


```{r}
null_distribution %>%
  visualise() + 
  shade_p_value(obs_stat = observed_stat, direction = "right")
```
```{r}
options(scipen = 999)
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "right")
p_value
```

## Difference in mean between paired samples. 

Can you say there is a row - wise association between groups?

```{r}
books <- read_csv("3_two_sample_tests/data/ucla_textbooks_f18.csv")
glimpse(books)
```
On average, is the price of 2-bedroom, 14-day holiday apartment lets in Florence significantly lower than that of comparable lets in Corfu?’

alpha = 0.05
H0: mu_diff = 0
Ha: mu_fiff != 0

```{r}
books_diff <- books %>%
  mutate(diff_new = bookstore_new - amazon_new) %>%
  filter(!is.na(diff_new))
```
```{r}
books_diff %>%
  ggplot(aes(x = diff_new)) + 
  geom_histogram(col = "white")
```

```{r}
null_distribution <- books_diff %>%
  specify(response = diff_new) %>%
  hypothesise(null = "point", mu = 0) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")
null_distribution

observed_stat <- books_diff %>%
  specify(response = diff_new) %>%
  calculate(stat = "mean")
observed_stat
```
```{r}
null_distribution %>%
  visualise() + 
  shade_p_value(obs_stat = observed_stat, direction = "both")
```
```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "both")
p_value
```

# Difference in proportions.

We would like to test whether the proportion of Obama voters who were women and the proportion of Romney voters who were women is significantly different (at a significance level 
alpha=0.05).

H0: women for Obama - women for Romney = 0
Ha: women for Obama - women for Romney != 0

```{r}
poll <- read_csv("3_two_sample_tests/data/election_data.csv")
glimpse(poll)
```
```{r}
ggplot(poll, aes(x = vote, fill = gender)) + 
  geom_bar(position = "fill") + 
  coord_flip()
```
```{r}
poll_observed_stat <- poll %>%
  specify(gender ~ vote, success = "woman") %>%
  calculate(stat = "diff in props", order = c("obama", "romney"))
poll_observed_stat
```
```{r}
poll_null_distribution <- poll %>%
  specify(gender ~ vote, success = "woman") %>%
  hypothesise(null = "independence") %>%
  generate(reps = 5000, type = "permute") %>%
  calculate(stat = "diff in props", order = c("obama", "romney"))
poll_null_distribution
```

```{r}
poll_null_distribution %>%
  visualise() + 
  shade_p_value(obs_stat = poll_observed_stat, direction = "both")
```
```{r}
p_value <- poll_null_distribution %>%
  get_p_value(obs_stat = poll_observed_stat, direction = "both")
p_value
```

# Task:

Calculate the 95% CI for the point estimate of the difference in the two proportions.

```{r}
bootstrap_distribution <- poll %>%
  specify(gender ~ vote, success = "woman") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "diff in props", order = c("obama", "romney"))
bootstrap_distribution
```
```{r}
bootstrap_prop_ci <- bootstrap_distribution %>%
  get_ci(level = 0.95, type = "percentile")
bootstrap_prop_ci
```
```{r}
bootstrap_distribution %>%
  visualise(bins = 30) + 
  shade_ci(endpoints = bootstrap_prop_ci)
```







