---
title: "Reading Data Lab"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE)
```

# Flat Files

Read in the following datasets. You may need define the variable types, skip rows or give your own variable names in each of them.

Before you read in each file, either view it in the Terminal (`less`) or open it in a text editor, so you can see what type it is. 

1. `user-for-x.txt`
2. `female_names.txt`
3. `edinburgh-trees.csv`

```{r}
library(readr)
uber_for_x <- read_tsv("data/uber-for-x.txt")
```
```{r}
uber_for_x
```


```{r}
female_names <- read_lines("data/female_names.txt")
```
```{r}
female_names
```


```{r}
edinburgh_trees <- read_csv("data/edinburgh-trees.csv", skip = 9, col_names = c("Latin Name", "Common Name", "Location", "Lat/Lon", "Age"))
```
```{r}
edinburgh_trees
```


```{r}
# You could read the names as a separate data-frame and use that to name the full data. Something like this:
tree_names <- read_csv("data/edinburgh-trees.csv", skip = 2, n_max = 5, col_names = c("letter", "name"))

trees <- read_csv("data/edinburgh-trees.csv", skip = 9, col_names = tree_names$name)
```
```{r}
trees
```

# Excel Files


Take the `populism` dataset that we read in earlier and write it to an Excel file. In the Excel file, you want the data for each region in a different sheet.


```{r, eval = FALSE}
library(readxl)
library(openxlsx)

populism <- read_excel("data/populism.xlsx")

populism_list <- split(populism, populism$region)

write.xlsx(populism_list, file='data/populism_by_region.xlsx')
```

# JSON

Parse the file `starwars.json` as:

1. a `list`
2. a `data.frame`


```{r}
library(jsonlite)

raw_starwars <- read_file("data/starwars.json")
starwars <- parse_json(raw_starwars)
```


```{r}
starwars_df <- fromJSON(raw_starwars)
```



